The TMB R package

### 2018-07-04 content/blog/2013-04-28-cross-referencing-reviewer-replies-in-latex.md

---
title: Cross-Referencing Line Numbers in LaTeX
subtitle: How to make writing reviewer response documents easier.
date: '2013-04-28'
slug: cross-referencing-reviewer-replies-in-latex
---

### 2018-07-04 content/blog/2013-12-01-plyr.Rmd
Merge hugo site

---
layout: post
title: "plyr: Split-Apply-Combine for Mortals"
subtitle: An introduction to the plyr R package.
published: true
---

@@

```{r d_ply_plot, fig.width=6, fig.height=2.5, fig.cap=""}
par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
d_ply(d, "year", transform, plot(count, main = unique(year), type = "o"))
mtext("count", side = 1, outer = TRUE, line = 1)
mtext("frequency", side = 2, outer = TRUE, line = 1)
```

>>>>

---
title: 'plyr: Split-Apply-Combine for Mortals'
subtitle: An introduction to the plyr R package.
date: '2013-12-01'
slug: plyr
---

@@

```{r d_ply_plot, fig.width=6, fig.height=2.5, fig.cap="", tidy=FALSE}
par(mfrow = c(1, 3), mar = c(2, 2, 1, 1), oma = c(3, 3, 0, 0))
d_ply(d, "year", transform, plot(count, main = unique(year), type = "o"))
mtext("count", side = 1, outer = TRUE, line = 1)
mtext("frequency", side = 2, outer = TRUE, line = 1)
```

### 2018-07-04 content/blog/2014-05-18-gamma-hurdle.Rmd
Merge hugo site

---
layout: post
title: Gamma Hurdle Models
subtitle: Fitting and interpreting Gamma hurdle models
published: true
---

@@

```{r hurdle1}
set.seed(1)
x <- 1:150
y <- rbinom(length(x), size = 1, prob = 0.7)
y <- y * rgamma(length(x), shape = 0.4)
non_zero <- ifelse(y > 0, 1, 0)
d <- data.frame(x, y, non_zero)
head(d)
library(ggplot2)
p <- ggplot(d, aes(x, y, colour = as.factor(non_zero))) + geom_point()
print(p)
```

>>>>

---
title: Gamma Hurdle Models
subtitle: Fitting and interpreting Gamma hurdle models
date: '2014-05-18'
slug: gamma-hurdle
---

@@

```{r hurdle1, cache=FALSE}
set.seed(1)
x <- 1:150
y <- rbinom(length(x), size = 1, prob = 0.7)
y <- y * rgamma(length(x), shape = 0.4)
non_zero <- ifelse(y > 0, 1, 0)
d <- data.frame(x, y, non_zero)
head(d)
library(ggplot2)
p <- ggplot(d, aes(x, y, colour = as.factor(non_zero))) + geom_point()
print(p)
```

### 2018-07-04 content/blog/2014-10-17-tmb.Rmd
Merge hugo site

---
layout: post
title: The TMB R package
subtitle: Life beyond ADMB?
published: true
---

@@

```{r, message=FALSE, results='hide', warning=FALSE}
obj <- MakeADFun(
  data = list(x = x, y = y), 
  parameters = list(a = 0, b = 0, log_sigma = 0),
  DLL = "regression")
opt <- do.call("optim", obj)
rep <- sdreport(obj)
```

@@

Which is about the same as we'd get from `stats::lm()`:

@@

Now let's try something a bit more exciting — a state space Gompertz population dynamics model. We'll have TMB integrate out the unobserved "true" state that is masked by observation error. You can represent the Gompertz model as:

\\[\begin{aligned}
y_t &= \ln N\_t\\\      
U\_t &amp;= a + b U\_{t-1} + \epsilon\_t\\\
\epsilon_t &amp;\sim \mathrm{Normal}(0, \sigma^2\_\mathrm{proc})\\\
y\_t &amp;\sim \mathrm{Normal}(U\_t, \sigma^2\_\mathrm{obs}),
\end{aligned}\\]

where \\(y\_t\\) is the observed \\(\ln\\) abundance (\\(N\\)) at time \\(t\\). The model is density independent if \\(b = 1\\), maximally density dependent if \\(b = 0\\), and inversely density dependence if \\(b < 0\\). The parameter \\(a\\) represents the expected ln abundance at the next time step when \\(y\_t = 0\\). The process noise is modelled as a normal distribution with mean of \\(0\\), and a standard deviation of \\(\sigma\_\mathrm{proc}\\). \\(U\\) represents the unobserved state vector, and \\(\sigma\_\mathrm{obs}\\) represents the standard deviation of observation error (on a log scale).

The TMB model for this state-space Gompertz model is:
```{r tmb-model}
tmb_model <- "
// State-space Gompertz model
#include <TMB.hpp>

@@

  data <- list(y = y)
  parameters <- list(a = 1, b = 0.5, log_sigma_proc = -1,
    log_sigma_obs = -1, u = rep(mean(y), N))
  obj <- MakeADFun(data, parameters, random = "u", DLL = "gompertztmb")
  obj$hessian <- FALSE
  opt <- do.call("optim", obj)
  rep <- sdreport(obj)

@@

(We just modelled a 5000 time step state space model in a couple of seconds!)

@@

```{r tmb-short, results='hide', message=FALSE, warning=FALSE, fig.height=8}
sim_gomp(N = 50)
```

>>>>

---
title: The TMB R package
subtitle: Life beyond ADMB?
date: '2014-10-17'
slug: tmb
---

```{r}
knitr::opts_chunk$set(cache = FALSE, autodep = TRUE)
```

@@

```{r, message=FALSE, results='hide', warning=FALSE}
obj <- MakeADFun(
  data = list(x = x, y = y),
  parameters = list(a = 0, b = 0, log_sigma = 0),
  DLL = "regression")
opt <- nlminb(start = obj$par, obj = obj$fn, gr = obj$gr)
rep <- sdreport(obj)
```

@@

Which is about the same as we'd get from `lm()`:

@@

Now let's try something a bit more exciting --- a state space Gompertz population dynamics model. We'll have TMB integrate out the unobserved "true" state that is masked by observation error. You can represent the Gompertz model as:

$$
\begin{aligned}
y_t &= \ln N_t\\
U_t &= a + b U_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Normal}(0, \sigma^2_\mathrm{proc})\\\
y_t &\sim \mathrm{Normal}(U_t, \sigma^2_\mathrm{obs}),
\end{aligned}
$$

where $y_t$ is the observed $\ln$ abundance ($N$) at time $t$. The model is density independent if $b = 1$, maximally density dependent if $b = 0$, and inversely density dependence if $b < 0$. The parameter $a$ represents the expected ln abundance at the next time step when $y_t = 0$. The process noise is modelled as a normal distribution with mean of $0$, and a standard deviation of $\sigma_\mathrm{proc}$. $U$ represents the unobserved state vector, and $\sigma_\mathrm{obs}$ represents the standard deviation of observation error (on a log scale).

The TMB model for this state-space Gompertz model is:

```{r tmb-model}
tmb_model <- "
// State-space Gompertz model
#include <TMB.hpp>

@@

  data <- list(y = y)
  parameters <- list(a = 1, b = 0.5, log_sigma_proc = -1,
    log_sigma_obs = -1, u = rep(mean(y), N))
  obj <- MakeADFun(data, parameters, random = "u", DLL = "gompertztmb")
  obj$hessian <- FALSE
  opt <- nlminb(start = obj$par, obj = obj$fn, gr = obj$gr)
  rep <- sdreport(obj)

@@

We just modelled a 5000 time step state space model in a couple seconds!

@@

```{r tmb-short, results='hide', message=FALSE, warning=FALSE, fig.height=8, tidy=FALSE}
sim_gomp(N = 50)
```
